### 2024-07-07 doc/Thesis/chapter/020-theory-materials.tex
edit until incl FTIR/XRD

%\pagebreak[4]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   SEM    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scanning Electron Microscopy}
%
\ds{
\td{FOR ME PERSONALLY TOO MUCH GENERAL INFORMATION}
The history of \gls{sem} can be traced back to 1843 when Scottish clockmaker Alexander Bain filed a patent for dissecting an image by scanning. 
The first \gls{sem} was build in 1933, 
the first images from scanning electron beams were published in 1935
and the first \gls{sem} was marketed in 1965.
A more detailed history of \gls{sem} can be read in a open-to-read paper by McMullan from 1995\cite{McMullan1995}. 
}
%
\Gls{sem} is a technique which allows visualization of surfaces with features in the nanometer regime. 
While optical microscopes use visible light and optical lenses, \gls{sem} uses accelerated electron beams and electrostatic and electromagnetic lenses.
This allows the generation of much more detailed images due to the shorter wavelengths of electrons compared to photons of visible light\cite{Kaliva2020}.
The electron beam produces X-rays, elastically backscattered (primary) electrons, inelastic (secondary) electrons and Auger electrons in the examined material. 
Secondary electrons carry information to conclude morphology and topology of the sample, while X-rays can be used to identify elements present in the surface. 
Electrons originate from either a \gls{feg}, where a strong electrical field rips electrons from the bulk, or from thermionic guns where a filament (tungsten W or \ch{LaB6} (brighter and longer lasting but more expensive)) is heated until electrons are emitted. 
Electrons are then accelerated by a voltage of \SI{2}{\kilo\volt} to \SI{40}{\kilo\volt} and bundled into narrow beams by lenses\cite{Vernon2000}.
All micrographs in this work were made at 
%Scanning electron microscopy was performed at 
an acceleration voltage of 5 kV.
A high mean free path is needed for electrons to travel from the source to the sample (and to the detector). 
Thus, a very low pressure is needed inside the microscope. 
In this work \gls{sem} was used as a preliminary way of checking the quality of the deposited layer. 
% squeeze condition into this page 
\enlargethispage{\baselineskip}
\todo{specifications of instruments used: acceleration voltage of 5 kV ??and with an in-lens detector??}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\iffalse
\subsection{UV/Vis/NIR Spectroscopy}
%%% WHAT? 
\Gls{uv}/\gls{vis}/\gls{nir} spectroscopy is a molecular spectroscopic method using 
interactions of \gls{uv}/\gls{vis}/\gls{nir} light 
(wavelengths of \num{e-7} to \num{e-6}\m{}) with molecules\cite{Schwedt2008}.
\td{FORMATTING: In general, light is described as periodic \gls{em} wave 
which - in vacuum - moves with the speed of light ($c = \mps{299792458}$)\cite{gHobel2006international}.
In other materials light moves with the velocity $v=c/n(\lambda)$, where $n(\lambda)$
is the refractive index which is dependent on the wavelength $\lambda$. 
%is the refractive index which is a function of the wavelength $\lambda$. 
The relation of the energy of a photon~$E$, its frequency~$\nu$, wavelength~$\lambda$ and wave number~$\bar{\nu}$ are as follows:}
\begin{align*}
	E &= h \cdot \nu \\
	\nu &= c/ \lambda \\
	\bar{\nu} &= 1/\lambda,
\end{align*}
where $h$ is Planck's constant.
%In words: shorter wavelength (higher frequency) photons are more energetic.
In practice the spectrum of \gls{em} waves is sectioned into different ranges; % (\td{see figure?+table});
from high to low energy: X-ray, \gls{uv}, \gls{vis}, (near, middle and far) \gls{ir}, microwaves and radio waves. 
X-rays interact with core electrons, \gls{uv}/\gls{vis} with valence electrons, \gls{ir} with 
%\td{it's adsorbed by certain groups?)} 
vibrations of certain groups, microwaves with molecular rotations and radio with electron and nuclear spins. 
%
% UVVis
Two prerequisites need to be fulfilled in order for a \gls{uv}/\gls{vis} photon to be 
absorbed by a molecule or atom.
Firstly, the energy of the photon must be the same as the difference between two electronic states. %, a lower populated and a higher empty orbital.}
The electron will then be excited into a energetically higher orbital.
Secondly, the electronic spin state of the molecule has to be the same before and after, otherwise the transition is quantum mechanically "forbidden" and the absorption is extremely unlikely. 
\fi

% IR
\Gls{ir} light interacts with molecular vibrations, which is only a part of the degrees of freedom of a molecule.
A molecule has $F=3N$ ($N$ number of atoms) degrees of freedom, including translational $F_T=3$ and rotational $F_R=3$ (2 for linear molecules) movements. 
%
The number of vibrations can therefore be calculated to be
\begin{align*}
	F_V &= 3N - F_T - F_R = 3N - 6 \\
	F_V &= 3N - F_T - F_R = 3N - 5 \textrm{ (for linear molecules)}.
\end{align*}
Vibrations are classified in valence vibrations (change of bond length) and deformation (change of bond angle) vibrations\cite{Melker2006}. 
Only vibrations that change the dipole moment of the molecule are \gls{ir} active, except Raman spectroscopy. 
%\td(does Raman not also use IR)}
The absorption of \Gls{nir} electromagnetic waves involves excitations of non-fundamental vibrations, overtones and combination modes which lead to intricate patterns in the so called fingerprint region. 
These transition are mostly "forbidden" and allow analysis of bulk matter without extensive preparation
due to the deep penetration depth of \gls{nir} light\cite{bec2019breakthrough}.

%\td{describe simple IR with monochromator}

\subsubsection{Fourier Transform Spectrometer}
 %A ceramic material (Nernst glower) is heated to around \oc{1600} as light source. 
In the classic two-beam-spectrometer the light emitted from the light source is split, 
sent through the sample and the reference, and one of the two beams is alternately 
sent through a monochromator to a detector (often a thermopile).
%
In the \gls{ft} spectrometer the beam is sent through the sample, split and 
reflected from a static and from a moving mirror, recombined and detected by a photo 
multiplier (a device which transforms photons into electrical signals). 
How the two beams will interfere upon recombination depends on the optical path difference (also called retardation) of the two light beams\cite{Schwedt2008}.
In a \gls{ft} spectrometer the reference has to be measured before the sample.
%
%When the refractive index of two layers differ, light can be used to measure the thickness\cite{Dumin1967} of the optical less dense material.

%\td{Frank-Condon rules, spin verbot and symmetrie verbot are ausschlaggebend für the resulting spectrum. }
%differences to \gls{uv}\gls{vis}: transmission instead of absorbance, higher energy left instead right in plot, base line on top instead of bottom.

@@

>>>>

%\pagebreak[4]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   SEM    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scanning Electron Microscopy}
%
\ds{
\td{FOR ME PERSONALLY TOO MUCH GENERAL INFORMATION}
The history of \gls{sem} can be traced back to 1843 when Scottish clockmaker Alexander Bain filed a patent for dissecting an image by scanning. 
The first \gls{sem} was build in 1933, 
the first images from scanning electron beams were published in 1935
and the first \gls{sem} was marketed in 1965.
A more detailed history of \gls{sem} can be read in a open-to-read paper by McMullan from 1995\cite{McMullan1995}. 
}
%
\Gls{sem} is a technique which allows visualization of surfaces with features in the nanometer regime. 
While optical microscopes use visible light and optical lenses, \gls{sem} uses accelerated electron beams and electrostatic and electromagnetic lenses.
This allows the generation of much more detailed images due to the shorter wavelengths of electrons compared to photons of visible light\cite{Kaliva2020}.
The electron beam produces X-rays, elastically backscattered (primary) electrons, inelastic (secondary) electrons and Auger electrons in the examined material. 
Secondary electrons carry information to conclude morphology and topology of the sample, while X-rays can be used to identify elements present in the surface. 
Electrons originate from either a \gls{feg}, where a strong electrical field rips electrons from the bulk, or from thermionic guns where a filament (tungsten W or \ch{LaB6} (brighter and longer lasting but more expensive)) is heated until electrons are emitted. 
Electrons are then accelerated by a voltage of \SI{2}{\kilo\volt} to \SI{40}{\kilo\volt} and bundled into narrow beams by lenses\cite{Vernon2000}.
%%Scanning electron microscopy was performed at 
%All micrographs in this work were made at 
%an acceleration voltage of 5 kV.
All SEM micrographs were taken with a Zeiss Supra 40 with 5 kV acceleration voltages.
A high mean free path is needed for electrons to travel from the source to the sample (and to the detector). 
Thus, a very low pressure is needed inside the microscope. 
In this work \gls{sem} was used as a preliminary way of checking the quality of the deposited layer. 
% squeeze condition into this page 
\enlargethispage{\baselineskip}
%\todo{specifications of instruments used: acceleration voltage of 5 kV ??and with an in-lens detector??}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\iffalse
\subsection{UV/Vis/NIR Spectroscopy}
%%% WHAT? 
\Gls{uv}/\gls{vis}/\gls{nir} spectroscopy is a molecular spectroscopic method using 
interactions of \gls{uv}/\gls{vis}/\gls{nir} light 
(wavelengths of \num{e-7} to \num{e-6}\m{}) with molecules\cite{Schwedt2008}.
\td{FORMATTING: In general, light is described as periodic \gls{em} wave 
which - in vacuum - moves with the speed of light ($c = \mps{299792458}$)\cite{gHobel2006international}.
In other materials light moves with the velocity $v=c/n(\lambda)$, where $n(\lambda)$
is the refractive index which is dependent on the wavelength $\lambda$. 
%is the refractive index which is a function of the wavelength $\lambda$. 
The relation of the energy of a photon~$E$, its frequency~$\nu$, wavelength~$\lambda$ and wave number~$\bar{\nu}$ are as follows:}
\begin{align*}
	E &= h \cdot \nu \\
	\nu &= c/ \lambda \\
	\bar{\nu} &= 1/\lambda,
\end{align*}
where $h$ is Planck's constant.
%In words: shorter wavelength (higher frequency) photons are more energetic.
In practice the spectrum of \gls{em} waves is sectioned into different ranges; % (\td{see figure?+table});
from high to low energy: X-ray, \gls{uv}, \gls{vis}, (near, middle and far) \gls{ir}, microwaves and radio waves. 
X-rays interact with core electrons, \gls{uv}/\gls{vis} with valence electrons, \gls{ir} with 
%\td{it's adsorbed by certain groups?)} 
vibrations of certain groups, microwaves with molecular rotations and radio with electron and nuclear spins. 
%
% UVVis
Two prerequisites need to be fulfilled in order for a \gls{uv}/\gls{vis} photon to be 
absorbed by a molecule or atom.
Firstly, the energy of the photon must be the same as the difference between two electronic states. %, a lower populated and a higher empty orbital.}
The electron will then be excited into a energetically higher orbital.
Secondly, the electronic spin state of the molecule has to be the same before and after, otherwise the transition is quantum mechanically "forbidden" and the absorption is extremely unlikely.

\subsubsection{Infrared spectroscopy}
% IR
\Gls{ir} light interacts with molecular vibrations, which is only a part of the degrees of freedom of a molecule.
A molecule has $F=3N$ ($N$ number of atoms) degrees of freedom, including translational $F_T=3$ and rotational $F_R=3$ (2 for linear molecules) movements. 
%
The number of vibrations can therefore be calculated to be
\begin{align*}
	F_V &= 3N - F_T - F_R = 3N - 6 \\
	F_V &= 3N - F_T - F_R = 3N - 5 \textrm{ (for linear molecules)}.
\end{align*}
Vibrations are classified in valence vibrations (change of bond length) and deformation (change of bond angle) vibrations\cite{Melker2006}. 
Only vibrations that change the dipole moment of the molecule are \gls{ir} active, except Raman spectroscopy. 
%\td(does Raman not also use IR)}
The absorption of \Gls{nir} electromagnetic waves involves excitations of non-fundamental vibrations, overtones and combination modes which lead to intricate patterns in the so called fingerprint region. 
These transition are mostly "forbidden" and allow analysis of bulk matter without extensive preparation
due to the deep penetration depth of \gls{nir} light\cite{bec2019breakthrough}.
\fi
%\td{describe simple IR with monochromator}

\subsubsection{Fourier Transform Spectrometer}
 %A ceramic material (Nernst glower) is heated to around \oc{1600} as light source. 
In the classic two-beam-spectrometer the light emitted from the light source is split, 
sent through the sample and the reference, and one of the two beams is alternately 
sent through a monochromator to a detector (often a thermopile).
%
In the \gls{ft} spectrometer, on the other hand, the beam is sent through the sample, split and 
reflected from a static and from a moving mirror, recombined and detected by a photo 
multiplier (a device which transforms photons into electrical signals). 
%How the two beams will interfere upon recombination 
The interference of the two beams 
depends on the optical path difference (also called retardation) of the two light beams\cite{Schwedt2008}.
In a \gls{ft} spectrometer the reference has to be measured before the sample.

\Gls{ft}\gls{uv}/\gls{vis}/\gls{nir} transmittance and reflectance spectra (at \SI{20}{\degree} incident angle) were recorded with a Bruker Vertex 70 spectrometer with a quartz beam splitter, \SI{0.5}{\milli\meter} aperture and Gallium-Phosphide detector for ultra violet light (\SI{303}{\nano\meter}--\SI{588} {\nano\meter}) and a silicon detector for visual and \gls{nir} light (\SI{500}{\nano\meter}--\SI{1.2}{\micro\meter}). 
% gallium phosphide http://dx.doi.org/10.1051/epjconf/20134800028
For transmittance the light entered the sample from the side with the layer. 
The \gls{uv} and \gls{vis}/\gls{nir} spectra were merged in Opus software. % included in the spectrometer. 
%
%When the refractive index of two layers differ, light can be used to measure the thickness\cite{Dumin1967} of the optical less dense material.

%\td{Frank-Condon rules, spin verbot and symmetrie verbot are ausschlaggebend für the resulting spectrum. }
%differences to \gls{uv}\gls{vis}: transmission instead of absorbance, higher energy left instead right in plot, base line on top instead of bottom.
\td{specifications of instruments}

@@

\Gls{xrd} diffractograms were obtained with a Thermo Scientific ARL Equinox 100 X-Ray Diffractometer.
The diffractograms were measured for \SI{2}{\minute} in reflection mode with \ch{Cu}-$K_{\alpha}$ radiation ($\lambda$=\SI{1.541874}{\angstrom}).
%. (where are the conditions?)}
All \gls{xrd} diffractograms were taken at \SI{5}{\degree} incident angle and compared to the internal database.

### 2024-07-07 doc/Thesis/chapter/03-experimental.tex
edit until incl FTIR/XRD

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHARACTERISATION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterization of \ch{ZrO2} layers}
All characterization procedures were executed at room temperature.
All \gls{sem} micrographs were taken with a Zeiss Supra 40 with  \SI{5}{\kilo\volt} acceleration voltages.
%\td{5kV, raum temperature, vacuum???}
\Gls{ft}\gls{uv}/\gls{vis}/\gls{nir} transmittance and reflectance spectra (at \SI{20}{\degree} incident 
angle) were recorded with a Bruker Vertex 70 
%transmission [\%] and reflection?  which angle? 
spectrometer with a quartz beam splitter, \SI{0.5}{\milli\meter} aperture and 
Gallium-Phosphide detector for ultra violet light (\SI{303}{\nano\meter}--\SI{588}
% gallium phosphide http://dx.doi.org/10.1051/epjconf/20134800028
{\nano\meter}) and a silicon detector for visual and \gls{nir} light 
(\SI{500}{\nano\meter}--\SI{1.2}{\micro\meter}). For transmittance the light
entered the sample from the side with the layer. The \gls{uv} and \gls{vis}/\gls{nir} spectra were merged 
in Opus software. % included in the spectrometer. 
%\td{
\Gls{xrd} diffractograms were obtained with a Thermo Scientific ARL Equinox 100 X-Ray Diffractometer.
The diffractograms were measured for \SI{2}{\minute} in reflection mode with \ch{Cu}-$K_{\alpha}$ radiation ($\lambda$=\SI{1.541874}{\angstrom}).
%. (where are the conditions?)}
All \gls{xrd} diffractograms were taken at \SI{5}{\degree} incident angle and compared to the internal database.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHARACTERISATION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterization of \ch{ZrO2} layers}
All characterization procedures were executed at room temperature.
\iffalse
%All \gls{sem} micrographs were taken with a Zeiss Supra 40 with  \SI{5}{\kilo\volt} acceleration voltages.
%\td{5kV, raum temperature, vacuum???}
%
\Gls{ft}\gls{uv}/\gls{vis}/\gls{nir} transmittance and reflectance spectra (at \SI{20}{\degree} incident angle) were recorded with a Bruker Vertex 70 spectrometer with a quartz beam splitter, \SI{0.5}{\milli\meter} aperture and Gallium-Phosphide detector for ultra violet light (\SI{303}{\nano\meter}--\SI{588} {\nano\meter}) and a silicon detector for visual and \gls{nir} light (\SI{500}{\nano\meter}--\SI{1.2}{\micro\meter}). 
% gallium phosphide http://dx.doi.org/10.1051/epjconf/20134800028
For transmittance the light entered the sample from the side with the layer. 
The \gls{uv} and \gls{vis}/\gls{nir} spectra were merged in Opus software. % included in the spectrometer. 
%\td{
%
\Gls{xrd} diffractograms were obtained with a Thermo Scientific ARL Equinox 100 X-Ray Diffractometer.
The diffractograms were measured for \SI{2}{\minute} in reflection mode with \ch{Cu}-$K_{\alpha}$ radiation ($\lambda$=\SI{1.541874}{\angstrom}).
%. (where are the conditions?)}
All \gls{xrd} diffractograms were taken at \SI{5}{\degree} incident angle and compared to the internal database.
\fi
